Empirical Bayes analysis of \RNAseq{} data for detection of gene expression heterosis

### 2015-08-11 ebayes_rnaseq_heterosis.tex
blinded; updated date

\begin{abstract}
Heterosis, or hybrid vigor, refers to the enhancements in the phenotype of hybrid progeny relative to their inbred parents. Although the heterosis phenomenon is extensively utilized in agriculture, the molecular basis of heterosis is still largely unknown. In an effort to understand phenotypic heterosis at the molecular level, researchers have begun to measure transcript abundance levels of thousands of genes in parental inbred lines and their hybrid offspring using RNA sequencing (\RNAseq{}) technology.  The resulting data allow researchers to search for evidence of gene expression heterosis as one potential molecular mechanism underlying heterosis of agriculturally important traits.  The null hypotheses of greatest interest in testing for gene expression heterosis are composite null hypotheses that are difficult to test with standard statistical approaches for \RNAseq{} analysis. To address these shortcomings, we develop a hierarchical negative binomial model and draw inferences using a computationally tractable empirical Bayes approach to inference. We demonstrate improvements over alternative methods via a simulation study based on a maize experiment and then analyze that maize experiment with our newly proposed methodology. 
% This article has supplementary material online.
\end{abstract}

@@

Notwithstanding these improvements, we believe our approach is a computationally efficient method that can immediately aid scientists who are interested in identifying candidate genes to explore a genetic mechanism of heterosis.

@@

%\section*{Supplementary Materials}

%The supplementary materials include three plain text files: the maize experiment data ({\tt data.csv}), a script to run our method ({\tt script.R}) on this data, and the \Stan{} model ({\tt model.stan}).

>>>>

\begin{abstract}
Heterosis, or hybrid vigor, refers to the enhancements in the phenotype of hybrid progeny relative to their inbred parents. Although the heterosis phenomenon is extensively utilized in agriculture, the molecular basis of heterosis is still largely unknown. In an effort to understand phenotypic heterosis at the molecular level, researchers have begun to measure transcript abundance levels of thousands of genes in parental inbred lines and their hybrid offspring using RNA sequencing (\RNAseq{}) technology.  The resulting data allow researchers to search for evidence of gene expression heterosis as one potential molecular mechanism underlying heterosis of agriculturally important traits.  The null hypotheses of greatest interest in testing for gene expression heterosis are composite null hypotheses that are difficult to test with standard statistical approaches for \RNAseq{} analysis. To address these shortcomings, we develop a hierarchical negative binomial model and draw inferences using a computationally tractable empirical Bayes approach to inference. We demonstrate improvements over alternative methods via a simulation study based on a maize experiment and then analyze that maize experiment with our newly proposed methodology. This article has supplementary material online.
\end{abstract}

@@

Notwithstanding these improvements, we believe our approach is a computationally efficient method that can immediately aid scientists who are interested in identifying candidate genes for a genetic mechanism of heterosis.

@@

\section*{Supplementary Materials}

The supplementary materials include these plain text files: the maize experiment data ({\tt data.csv}), a script to run our method ({\tt script.R}) on these data, a script to create the figures ({\tt figs.R}), two functions ({\tt get\_hyperparameters.R} and {\tt single\_gene\_analysis.R}), and the \Stan{} model ({\tt model.stan}).
